Verifying Spatial Properties of Array Computations

### 2017-04-18 oopsla17/evaluation.tex
refactor example

\begin{minted}[breakafter=+:,breakindent=-0.6em,breaksymbolsep=0.4em,linenos=false,xleftmargin=-0.5em]{fortran}
  !=stencil readOnce, (pointed(dim=1))*(pointed(dim=2))*(pointed(dim=3))
\end{minted}

@@

\begin{minted}[breakafter=+:,breakindent=-0.6em,breaksymbolsep=0.4em,linenos=false,xleftmargin=-0.5em]{fortran}
  !=stencil readOnce, (forward(depth=1, dim=3, nonpointed))*(pointed(dim=1))*(pointed(dim=2))
\end{minted}

@@

One of the initial goals of CamFort was to provide tools that enable
optimisation and refactoring without changing behaviour, and the
stencil specification feature was designed with that in mind. But
while we were perusing real-world source bases, we realised that there
were also many cases where CamFort could help with refactoring
precisely because they required changes in behaviour. For example, a
relatively common change observed in the logs of revision control for
various projects is the refactoring of array indices. Either
re-ordering, adding or deleting indices. Any of these, if not
perfectly propagated throughout all uses, could result in unexpected
outcomes. However, CamFort will pick up the difference during its
stencil specification checking. Ideally, an experienced programmer may
take advantage of the region variable feature of CamFort in order to
minimise the number of specifications that need changing and gain the
full time-saving advantage. But even if not, a quick search for the
specifications can have them updated quickly and then any code that is
not updated will trigger an error when checked.

>>>>

\begin{minted}[breakafter=+:,breakindent=-0.6em,breaksymbolsep=0.4em,linenos=false,xleftmargin=-0.5em]{fortran}
  != stencil readOnce, pointed(dim=1)*pointed(dim=2)*pointed(dim=3)
\end{minted}

@@

\begin{minted}[breakafter=+:,breakindent=-0.6em,breaksymbolsep=0.4em,linenos=false,xleftmargin=-0.5em]{fortran}
  != stencil readOnce, forward(depth=1, dim=3, nonpointed)*pointed(dim=1)*pointed(dim=2)
\end{minted}

@@

One of the initial goals of CamFort was to provide tools that enable
optimisation and refactoring without changing behaviour, and the
stencil specification feature was designed with that in mind. But
while we were perusing real-world source bases, we realised that there
were also many cases where CamFort could help with refactoring
precisely because they required changes in behaviour. For example, a
relatively common change observed in the logs of revision control for
various projects is the refactoring of array dimensions. Either
re-ordering, adding or deleting dimensions. Any of these, if not
perfectly propagated throughout all uses, could result in unexpected
outcomes. However, CamFort will pick up the difference during its
stencil specification checking. Ideally, an experienced programmer may
take advantage of the region variable feature of CamFort in order to
minimise the number of specifications that need changing and gain the
full time-saving advantage. But even if not, a quick search for the
specifications can have them updated quickly and then any code that is
not updated will trigger an error when checked.

An example of a refactored region adding a single dimension to an array, adapted from the UM:
\begin{minted}[breakafter=+:,breakindent=-0.6em,breaksymbolsep=0.4em,linenos=false,xleftmargin=-0.5em]{fortran}
  ! before
  != region :: r = readOnce, pointed(dim=1)
  ! after
  != region :: r = readOnce, pointed(dim=1)*pointed(dim=2)
\end{minted}
